# Model Prediction

Required Packages

```{r Required Packages}

library(plyr)

```

Load models from Rdata file

```{r Load Models}

load("trained_models_NoSch.Rdata")
load("trained_pumps_on_models.Rdata")

hhpu_model <- trained_models$DForest
pumps_on_model <- PumpsOn_Models$DTree

```

Load the input data. Replace "Line2.xdf" with whatever the input file is. Data headers should be:
* ts_pi (timestamp)
* flow_rate
* dra
* pump_station
* line_segment
* ts_month
* ts_day
* ts_hour
* ts_dow

```{r Load Data}

Line2_Pred <- RxXdfData("Line2_Pred_Test.xdf")

```

Predict the number of pumps on. The input data for this model is:
* flow_rate (sim_rate).
* pump_station.

```{r Predict Pumps On}

keep_vars <- names(Line2_Pred)
Line2_Pred <- rxPredict(pumps_on_model, Line2_Pred, outData = "Line2_pumps_on_pred.xdf", extraVarsToWrite = keep_vars)
plyr::rename(pumps_on_pred, c("0_prob" = "prob_0", "1_prob" = "prob_1", "2_prob" = "prob_2", "3_prob" = "prob_3",
                              "4_prob" = "prob_4", "5_prob" = "prob_5", "6_prob" = "prob_6"))

predict_PumpsOn <- function(dat){
  dat$pump_status_on <- round(0*dat$prob_0 + 1*dat$prob_1 + 2*dat$prob_2 + 3*dat$prob_3 + 
                               4*dat$prob_4 + 5*dat$prob_5 + 6*dat$prob_6)
  dat$pump_status_on <- factor(dat$pump_status_on, levels = c(as.character(0:6)))
  return(dat)
}

rxDataStep(Line2_Pred, outFile = "Line2_pumps_on_pred1.xdf", transformFunc = predict_PumpsOn)
rxDataStep("Line2_pumps_on_pred1.xdf", outFile = "Line2_pumps_on_pred2.xdf",
           varsToDrop = c("prob_0", "prob_1", "prob_2", "prob_3", "prob_4", "prob_5", "prob_6"))
file.remove(c("Line2_pumps_on_pred.xdf", "Line2_pumps_on_pred1.xdf"))
file.rename("Line2_pumps_on_pred2.xdf", "Line2_pumps_on_pred.xdf")

```

The header outfile is:
* ts_pi (timestamp)
* flow_rate
* dra
* pump_station
* line_segment
* ts_month
* ts_day
* ts_hour
* ts_dow
* pump_total
* pump_status_on

Now to predict hhpu. The previous header output is this chunks header input. We can calculate KW here as well.

```{r Predict HHPU}

keep_vars <- names(Line2_Pred)
rxPredict(hhpu_model, Line2_Pred, outData = "Line2_pumps_on_pred1.xdf", extraVarsToWrite = keep_vars)
rxDataStep("Line2_pumps_on_pred1.xdf", outFile = "Line2_pumps_on_pred2.xdf", transforms = list(kw_Pred = hhpu_Pred*0.7457))
file.remove(c("Line2_pumps_on_pred.xdf", "Line2_pumps_on_pred1.xdf"))
file.rename("Line2_pumps_on_pred2.xdf", "Line2_pumps_on_pred.xdf")

```

The final header output is:
* ts_pi (timestamp)
* flow_rate
* dra
* pump_station
* line_segment
* ts_month
* ts_day
* ts_hour
* ts_dow
* pump_total
* pump_status_on
* hhpu_Pred
* kw_Pred

