# Part 3b - Data Exploration 1: Label Exploration

Required Packages

```{r Required Packages, echo=FALSE, warning=FALSE}

library(dplyrXdf)
library(ggplot2)

```

Data to be loaded. We created a new dataset during data_exploration_feature which added columns for pump station type as well as indicators for where each station lies along the line. We'll use that data set here

```{r Load Data}

Line2 <- RxXdfData("Line2_Seq.xdf")

```

Let's turn our attention to our label hhpu

Distribution of HHPU

```{r HHPU Distribution}

rxHistogram(~hhpu,Line2)

```

A lot of hhpu = 0 points. Let's calculate the percentage of data points sucht that hhpu = 0

```{r HHPU = 0 Porportion}

nhhpu0 <- Line2 %>% filter(hhpu == 0) %>% nrow
phhpu0 <- nhhpu0/nrow(Line2)*100
cat("the percentage of points where hhpu = 0 is ", round(phhpu0,2),"%",sep="")

```

Let's look at the dist of hhpu by pump_station

```{r HHPU dist by NumPumpsOn, echo=FALSE}

plt <- rxHistogram(~hhpu|pump_station,Line2)
for(i in 1:29){
  plot(plt[i])
}

```

Can't see much so let's look at hhpu by segment

```{r HHPU by Segment}

rxHistogram(~hhpu|line_segment,Line2)

```

Let's look at the distribution of hhpu for each segment and station

```{r HHPU Distribution by Segment and Station}

rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-EP-YP"), title = "2-EP-YP")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-YP-KB"), title = "2-YP-KB")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-KB-SB"), title = "2-KB-SB")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-QU-CM"), title = "2-QU-CM")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-CM-GF"), title = "2-CM-GF")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-GF-CR"), title = "2-GF-CR")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-CR-PR"), title = "2-CR-PR")

```

Segments 2-CM-CR and 2-CR-PR had Plenty of points where HHPU didn't equal 0. For the other 4 segments, the majority of points had hhpu = 0.

Let's calculate the percentage of points where hhpu = 0 for each segment.

```{r Proportion of Points HHPU = 0 by segment, echo=FALSE}

hhpu0 <- rxDataStep(Line2, outFile = "Line2_hhpu0.xdf", rowSelection = (hhpu == 0), overwrite = T)

hhpu0_len <- rxCube(~line_segment, hhpu0, returnDataFrame = T)
hhpu_len <- rxCube(~line_segment, Line2, returnDataFrame = T)
hhpu0_len$Total_Counts <- hhpu_len$Counts
hhpu0_len$Proportion <- hhpu0_len$Counts/hhpu0_len$Total_Counts*100

print(hhpu0_len)

```

Similar to flow = 0. The first 4 segments saw a large amount of downtime (~40-50%) where as the last 3 segments saw little downtime roughly (5-10%).

Let's look at hhpu by pump_status_on

```{r HHPU by pump_status_on}

rxHistogram(~hhpu|pump_status_on,Line2)

```

Calculate the percentage of data points where NumPumpsOn = 0 given hhpu = 0

```{r pumps_status_on = 0 given hhpu = 0}

hhpu0n0 <- hhpu0 %>% filter(pump_status_on == 0) %>% nrow()
phhpu0n0 <- hhpu0n0/nrow(hhpu0)*100
cat("the percentage of points where pump_status_on = 0 given hhpu = 0 is ",round(phhpu0n0,2),"%",sep="")

```

Given hhpu = 0, every pump is off for about 99.5% of the pumps. Previously we found that 99.3% of points had flow = 0 given all the pumps were off.

Let's look at the percentage of data points where hhpu = 0 given num_pumps_on = 0

```{r Percentage and Dist hhpu = 0 given pump_status_on = 0}

pump0 <- rxDataStep(Line2, outFile = "pump0.xdf", rowSelection = (pump_status_on == 0), overwrite = T)
nhhpu0 <- pump0 %>% filter(hhpu == 0) %>% nrow()
phhpu0 <- nhhpu0/nrow(pump0)*100

cat("the percentage of points where hhpu = 0 given pump_status_on = 0 is ",round(phhpu0,2),"%",sep="")
rxHistogram(~hhpu,pump0)

```

Let's see how often flow = 0 given hhpu = 0.

```{r Percentage flow = 0 given hhpu = 0}

hhpu0f0 <- hhpu0 %>% filter(flow_rate == 0) %>% nrow()
phhpu0f0 <- hhpu0f0/nrow(hhpu0)*100
cat("the percentage of points where flow = 0 given hhpu = 0 is ",round(phhpu0f0,2),"%",sep="")

```

So roughly 66% of the data points have flow = 0 given hhpu = 0. This is probably because it takes some time for the liquid to stop flowing once the pumps are shut off. Let's determine the distribution of flow_rate given hhpu = 0

```{r Flow Distribution given hhpu = 0}

rxHistogram(~flow_rate,hhpu0)

```

Let's look hhpu = 0 given flow = 0 distribution and porportion.

```{r HHPU = 0 given Flow = 0}

flow0 <- rxDataStep(Line2, outFile = "flow0.xdf", rowSelection = (flow_rate == 0), overwrite = T)
flow0h0 <- flow0 %>% select(hhpu) %>% nrow()
pflow0h0 <- flow0h0/nrow(flow0)*100

cat("the percentage of points where hhpu = 0 given flow = 0 is ", round(pflow0h0,2),"%",sep="")
rxHistogram(~hhpu,flow0)

```

given flow = 0, approximately 100% of the points had hhpu = 0.

So while having no pumps on and hhpu = 0 will not always mean flow = 0 (it takes time to slow down), if flow = 0, hhpu and pump_status_on will almost always be 0.

# Removing HHPU = 0 from the data

We can see that the vast majority of points with hhpu = 0 also had pump_status_on = 0 but because of this we can't see much else so let's filter out the points = 0.

####NOTE TO SELF: perform data exploration on on a data set with hhpu = 0 filtered out

```{r HHPU != 0 by pump_status}

hhpu_n0 <- rxDataStep(Line2, outFile = "Line2_hhpu_not0.xdf", rowSelection = (hhpu != 0), overwrite = T)

rxHistogram(~hhpu,hhpu_n0)
rxHistogram(~hhpu|pump_status_on,hhpu_n0)

```

Clearly the number of pumps on increases the hhpu value.

## HHPU vs. Time variables

hhpu by day, day of the week, hour, and month

```{r HHPU by Datetime Distribution}

rxHistogram(~hhpu|ts_day, Line2, title = "Day")
rxHistogram(~hhpu|ts_dow, Line2, title = "Day of Week")
rxHistogram(~hhpu|ts_hour, Line2, title = "Hour")
rxHistogram(~hhpu|ts_month, Line2 , title = "Month")

```

HHPU = 0 is killing this so let's remove the data points where hhpu = 0

```{r HHPU != 0 by Datetime Distribution}

rxHistogram(~hhpu|ts_day, hhpu_n0, title = "Day")
rxHistogram(~hhpu|ts_dow, hhpu_n0, title = "Day of Week")
rxHistogram(~hhpu|ts_hour, hhpu_n0, title = "Hour")
rxHistogram(~hhpu|ts_month, hhpu_n0, title = "Month")

```

Month seems to be the datetime values which seem to matter.

Let's look at hhpu and the interactions between dow and the other datetime features.

HHPU by day of the week and hour.

```{r HHPU by dow and Hour}

rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Mon"), title = "Monday")
rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Tue"), title = "Tuesday")
rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Wed"), title = "Wednesday")
rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Thu"), title = "Thursday")
rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Fri"), title = "Friday")
rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Sat"), title = "Saturday")
rxHistogram(~hhpu|ts_hour, hhpu_n0, rowSelection = (ts_dow == "Sun"), title = "Sunday")

```

HHPU by day of the week and day.

```{r HHPU by dow and day}

rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Mon"), title = "Monday")
rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Tue"), title = "Tuesday")
rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Wed"), title = "Wednesday")
rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Thu"), title = "Thursday")
rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Fri"), title = "Friday")
rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Sat"), title = "Saturday")
rxHistogram(~hhpu|ts_day, hhpu_n0, rowSelection = (ts_dow == "Sun"), title = "Sunday")

```

HHPU by day of the week and month.

```{r HHPU by dow and month}

rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Mon"), title = "Monday")
rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Tue"), title = "Tuesday")
rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Wed"), title = "Wednesday")
rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Thu"), title = "Thursday")
rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Fri"), title = "Friday")
rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Sat"), title = "Saturday")
rxHistogram(~hhpu|ts_month, hhpu_n0, rowSelection = (ts_dow == "Sun"), title = "Sunday")

```

Let's look at hhpu wrt pump station type

```{r HHPU vs station type}

rxHistogram(~hhpu|Type, Line2)

```

The distribution doesn't seem to change but let's remove hhpu = 0 to be sure.

```{r HHPU vs station type}

rxHistogram(~hhpu|Type, Line2, rowSelection = (hhpu != 0))

```

Distribution does not seem to change all that much.