---
title: "Part 7a - Model Validation - New Pi Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Required Packages, warning=FALSE}

# Needed for importing data from R. Remove these if you've already imported the new data.
library(httr)
library(RCurl)
library(bitops)

# Packages needed for data cleansing.
library(plyr)
library(tidyr)
library(dplyrXdf)
library(lubridate)

# Packages needed for graphics
library(ggplot2)

```

Change XDF file function. See Part 2, raw data exploration for more details.

```{r Xdf File Function}

change_Xdf <- function(starting_filename = "Line2.xdf", final_filename = starting_filename, n = 1){
  
  xdf <- starting_filename
  xdf_name <- substr(xdf, 1, nchar(xdf) - 4)
  xdf_ext <- ".xdf"
  i <- 0
  
  while (i < n){
    if (i == 0){
      file.remove(xdf)
      i <- i + 1
    }
    else {
      xdf_str <- paste0(xdf_name,i,xdf_ext)
      file.remove(xdf_str)
      i <- i + 1
    }
  }
  
  xdf_str <- paste0(xdf_name,i,xdf_ext)
  file.rename(xdf_str, final_filename)
  return(RxXdfData(final_filename))
  
}

```

Import new pi data from blob storage. See part 1 for more details.

NOTE: Skip this step if you've already imported the new pi data.

```{r Import Data}

# Functions needed to call the blob storage and import new data.
azureStorageCall <- function(url, verb, key, storageType="file", requestBody=NULL, headers=NULL, ifMatch="", md5="") { 
  urlcomponents <- httr::parse_url(url)
  storageDomainName <- paste0(".", storageType, ".core.windows.net")
  account <- gsub(storageDomainName, "", urlcomponents$hostname, fixed = TRUE)
  container <- urlcomponents$path
  
  lct <- Sys.getlocale("LC_TIME")
  Sys.setlocale("LC_TIME", "en_US.UTF-8")
  `x-ms-date` <- format(Sys.time(),"%a, %d %b %Y %H:%M:%S %Z", tz="GMT")
  Sys.setlocale("LC_TIME", lct)
  
  `Content-Length` <- ""; `Content-Type` <- ""
  if(!is.null(requestBody)) {
    if(class(requestBody) == "form_file") {
      `Content-Length` <- (file.info(requestBody$path))$size
      `Content-Type` <- requestBody$type 
    } else {
      requestBody <- enc2utf8(as.character(requestBody))
      `Content-Length` <- nchar(requestBody, "bytes")
      `Content-Type` <- "text/plain; charset=UTF-8" 
    }
  } 
  
  headers <- setNames(c(`x-ms-date`, "2015-04-05",  unlist(headers)), 
                      c("x-ms-date", "x-ms-version", unclass(names(unlist(headers)))))
  headers <- headers[order(names(headers))]
  CanonicalizedHeaders <- paste(names(headers), headers, sep=":", collapse = "\n")
  
  if(!is.null(urlcomponents$query)) {
    components <- setNames(unlist(urlcomponents$query), unclass(names(unlist(urlcomponents$query))))
    componentstring <- paste0("\n", paste(names(components[order(names(components))]),
                                          components[order(names(components))], sep=":", collapse = "\n"))
  } else componentstring <- ""
  CanonicalizedResource <- paste0("/",account,"/",container, componentstring)
  
  signaturestring <- paste0(verb, "\n\n\n", `Content-Length`, "\n", md5, "\n", `Content-Type`, "\n\n\n", 
                            ifMatch, "\n\n\n\n", CanonicalizedHeaders, "\n", CanonicalizedResource)
  
  requestspecificencodedkey <- RCurl::base64(
    digest::hmac(key=RCurl::base64Decode(key, mode="raw"),
                 object=enc2utf8(signaturestring),
                 algo= "sha256", raw=TRUE)
  )
  
  authorizationtoken <- paste0("SharedKey ", account, ":", requestspecificencodedkey)
  
  headers_final <- add_headers(Authorization=authorizationtoken, headers, `Content-Type` = `Content-Type`)
  call <- httr::VERB(verb=verb, url=url, config=headers_final, body=requestBody, verbose())
  
  return(call)
} 

azureCSVBZip2ToDataTable <- function (azureStorageUrl, key, container, bzip2Filepath, storageType="file", encoding = 'UTF-8') {
  
  container <- gsub(" ", "%20", container)  
  bzip2Filepath <- gsub(" ", "%20", bzip2Filepath)  
  
  azureUrl <- paste0(azureStorageUrl, "/", container, "/", bzip2Filepath) 
  
  response <- azureStorageCall(azureUrl, "GET", key, storageType)
  
  bzip2 <- content(response, as="raw")
  
  tmpBzip2Filename <- tempfile(pattern = "file", tmpdir = tempdir(), fileext = ".csv.bz2")
  writeBin(bzip2, tmpBzip2Filename)
  
  remove(bzip2)
  
  con <- bzfile(tmpBzip2Filename)
  df <- read.csv(con, stringsAsFactors = F)

  file.remove(tmpBzip2Filename)  
  
  return (df)
}

# Key, Url, and Share strings needed to access blob storage
key <- "ITnD8P0dMGURFECkMANiwtS65Ou72OKPf1WtUYLFmUBs7sJW2UhghelMKrZkZxRAXu3b6aaVeqhFZey6MosRaw=="
azureBaseUrl <- "https://enbze2elpdestarserver.blob.core.windows.net"
azureShare <- "2017-power-optimization"

# Import Pi Data
azureZipFilename <- "machine_learning/data/2017/Line2_Month01_2017.csv.bz2"
Pi2 <- rxImport(inData = azureCSVBZip2ToDataTable(azureBaseUrl, key, azureShare, azureZipFilename, "blob",
                                                        encoding="UTF-8"),
                     outFile = "Line2_Validate_Raw.xdf", overwrite = T, stringsAsFactors = F)
azureZipFilename <- "machine_learning/data/2017/Line2_Month02_2017.csv.bz2"
Pi2 <- rxImport(inData = azureCSVBZip2ToDataTable(azureBaseUrl, key, azureShare, azureZipFilename, "blob",
                                                        encoding="UTF-8"),
                     outFile = "Line2_Validate_Raw.xdf", overwrite = T, append = T, stringsAsFactors = F)

```

Now we need to explore the raw data to make sure it is good and find what needs to be cleaned up. See Part 2 for more details.

```{r Raw Data Exploration}

# Convert certain column types to factors
tags <- c("HHPU", "KW", "Discharge_Pressure", "Flow_Rate", "Suction_Pressure", "DRA",
          "Pump_Status_1", "Pump_Status_2", "Pump_Status_3", "Pump_Status_4", "Pump_Status_5")

convert_Col <- function(dat){
  dat$LogTime <- as.POSIXct(dat$LogTime, format = "%Y-%m-%dT%H:%M:%S", tz = "MST")
  dat$Seg <- as.factor(dat$Seg)
  dat$Site <- as.factor(dat$Site)
  dat$TagType <- as.factor(dat$TagType)
  dat$ValType <- as.factor(dat$ValType)
  dat$IsGood <- as.factor(dat$IsGood)
  dat$IsQuestionable <- as.factor(dat$IsQuestionable)
  dat$IsSubst <- as.factor(dat$IsSubst)
  return(dat)
}

rxDataStep(inData = Pi2, outFile = "Line2_Validate_RawC.xdf", rowSelection = (TagType %in% t),
           transformObjects = list(t = tags), overwrite = TRUE)
rxDataStep("Line2_Validate_RawC.xdf", outFile = "Line2_Validate_RawC1.xdf", transformFunc = convert_Col)
Pi2 <- change_Xdf(starting_filename = "Line2_Validate_RawC.xdf", n = 1)

# Find Missing and bad Tags
rxCrossTabs(~TagType:Site, Pi2)
rxCrossTabs(~Site:IsGood:TagType, Pi2)

```

Only missing dra tags for WC and the only significant amount of bad tags was for dra tags.

Start Cleansing the raw data for modelling purposes. See Part 3 Cleansing Data for more details

```{r Cleansing Data}

# Remove Unecessary Columns and Tags
vars <- c("LogTime","Seg", "Site", "TagType", "ValType", "ValInt", "ValFlt", "ValDig", "IsGood", "LogMonth", "LogYear")
tags <- c("HHPU", "KW", "Discharge_Pressure", "Flow_Rate", "Suction_Pressure", "DRA",
          "Pump_Status_1", "Pump_Status_2", "Pump_Status_3", "Pump_Status_4", "Pump_Status_5")

Pi2 <- RxXdfData("Line2_Validate_Raw.xdf")
rxDataStep(inData = Pi2, outFile = "Line2_Validate.xdf", overwrite = T, varsToKeep = vars)
rxDataStep(inData = "Line2_Validate.xdf", outFile = "Line2_Validate1.xdf", rowSelection = (TagType %in% t),
           transformObjects = list(t = tags), rowsPerRead = 10000000)
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 1)

# Create Value Column
create_Value <- function(dat){
  dat$Value <- ifelse(dat$ValType == "int", as.character(dat$ValInt),
                      ifelse(dat$ValType == "float", as.character(dat$ValFlt), dat$ValDig))
  return(dat)
}

rxDataStep(Pi2, outFile = "Line2_Validate1.xdf", overwrite = T, transformFunc = create_Value)
rxDataStep("Line2_Validate1.xdf", outFile = "Line2_Validate2.xdf",
           varsToDrop = c("ValType", "ValInt", "ValFlt", "ValDig"))
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 2)

# Fix bad DRA values and filter out the other bad readings
fix_DRA_bad <- function(dat){
  dat$Value <- ifelse((dat$TagType == "DRA") & (dat$IsGood == 0), "0", dat$Value)
  dat$IsGood <- ifelse((dat$TagType == "DRA") & (dat$IsGood == 0), as.integer(1), as.integer(dat$IsGood))
  return(dat)
}

rxDataStep(Pi2, outFile = "Line2_Validate1.xdf", transformFunc = fix_DRA_bad)
rxDataStep(inData = "Line2_Validate1.xdf", outFile = "Line2_Validate2.xdf", rowSelection = (IsGood == 1))
rxDataStep(inData = "Line2_Validate2.xdf", outFile = "Line2_Validate3.xdf", varsToDrop = c("IsGood"))
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 3)

# Pivot Data by month
Pi2_df <- rxImport(Pi2, rowSelection = (LogMonth == 1))
Pi2_df <- spread(Pi2_df, TagType, Value)
Pi2_df <- Pi2_df[c(1,2,3,4,5,8,6,16,7,9,10,11,12,13,14,15)]
rxImport(Pi2_df, outFile = "Line2_Validate_AfterPivot.xdf")
Pi2_df <- rxImport(Pi2, rowSelection = (LogMonth %in% c(2,3)))
Pi2_df <- spread(Pi2_df, TagType, Value)
Pi2_df <- Pi2_df[c(1,2,3,4,5,8,6,16,7,9,10,11,12,13,14,15)]
rxImport(Pi2_df, outFile = "Line2_Validate_AfterPivot.xdf", overwrite = T, append = T)
file.rename("Line2_Validate.xdf", "Line2_Validate_UnPivot.xdf")
Pi2 <- RxXdfData("Line2_Validate_AfterPivot.xdf")

# Fix Missing DRA Values
fix_Missing <- function(dat){
  dat$DRA <- ifelse(is.na(dat$DRA) == TRUE, "0", dat$DRA)
  return(dat)
}

Pi2 <- rxDataStep(Pi2, outFile = "Line2_Validate.xdf", transformFunc = fix_Missing)

# Split Segment 2-CM-CR into 2-CM-GF and 2-GF-CR
split_Seg <- function(dat){
  dat$Seg <- ifelse(dat$Site %in% c("CM", "SP", "LP", "XP"), "2-CM-GF",
                    ifelse(dat$Site %in% c("GF", "PL", "DN", "VG"), "2-GF-CR", dat$Seg))
  return(dat)
}

rxDataStep(Pi2, outFile = "Line2_Validate1.xdf", transformFunc = split_Seg)
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 1)

# Remove inactive stations
rxDataStep(Pi2, outFile = "Line2_Validate1.xdf", rowSelection = !(Site %in% c("IV", "PW")))
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 1)

# Convert column types to their proper type. We're going to load data into memory and leave it there for the next step.
conv_Columns <- function(dat){
  dat$LogTime <- as.POSIXct(dat$LogTime, format = "%Y-%m-%dT%H:%M:%S", tz = "UTC")
  dat$Seg <- as.factor(dat$Seg)
  dat$Site <- as.factor(dat$Site)
  dat$LogMonth <- as.factor(dat$LogMonth)
  dat$LogYear <- as.factor(dat$LogYear)
  dat$HHPU <- as.numeric(dat$HHPU)
  dat$KW <- as.numeric(dat$KW)
  dat$Flow_Rate <- as.numeric(dat$Flow_Rate)
  dat$Discharge_Pressure <- as.numeric(dat$Discharge_Pressure)
  dat$Suction_Pressure <- as.numeric(dat$Suction_Pressure)
  dat$DRA <- as.numeric(dat$DRA)
  dat$Pump_Status_1 <- as.factor(dat$Pump_Status_1)
  dat$Pump_Status_2 <- as.factor(dat$Pump_Status_2)
  dat$Pump_Status_3 <- as.factor(dat$Pump_Status_3)
  dat$Pump_Status_4 <- as.factor(dat$Pump_Status_4)
  dat$Pump_Status_5 <- as.factor(dat$Pump_Status_5)
  return(dat)
}

Pi2_df <- rxImport(Pi2)
Pi2_df <- conv_Columns(Pi2_df)

# Create number of pumps on and total pumps feature
create_PumpVars <- function(dat, on_status = c("ON", "IN SEQUENCE ON")){
  
  dat$Pump_Status_N1 <- ifelse(dat$Pump_Status_1 %in% on_status, 1, 0)
  dat$Pump_Status_N2 <- ifelse(dat$Pump_Status_2 %in% on_status, 1, 0)
  dat$Pump_Status_N3 <- ifelse(dat$Pump_Status_3 %in% on_status, 1, 0)
  dat$Pump_Status_N4 <- ifelse(dat$Pump_Status_4 %in% on_status, 1, 0)
  dat$Pump_Status_N5 <- ifelse(dat$Pump_Status_5 %in% on_status, 1, 0)
  
  dat[c(17,18,19,20,21)][is.na(dat[c(17,18,19,20,21)])] <- 0
  dat$NumPumpsOn <- dat$Pump_Status_N1+dat$Pump_Status_N2+dat$Pump_Status_N3+dat$Pump_Status_N4+dat$Pump_Status_N5
  
  dat$TotalPumps <- ifelse(is.na(dat$Pump_Status_1) == TRUE, 0,
                           ifelse(is.na(dat$Pump_Status_2) == TRUE, 1,
                                  ifelse(is.na(dat$Pump_Status_3) == TRUE, 2,
                                         ifelse(is.na(dat$Pump_Status_4) == TRUE, 3,
                                                ifelse(is.na(dat$Pump_Status_5) == TRUE, 4, 5)))))
  
  
  dat <- dat[c(1:11,22,23)]
  return(dat)
}

#Pi2_df <- rxImport(Pi2)
Pi2_df <- create_PumpVars(Pi2_df, c("ON", "IN SEQUENCE ON"))
rxImport(Pi2_df, outFile = "Line2_Validate1.xdf", overwrite = T,
         colInfo = list(NumPumpsOn = list(type = "factor", levels = c(as.character(0:6))),
                        TotalPumps = list(type = "factor", levels = c(as.character(0:6)))))
remove(Pi2_df)
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 1)

# Create Datetime Variables
create_DatetimeVars <- function(dat){
  dat$LogTime <- with_tz(dat$LogTime, tzone = "MST")
  dat$Month <- factor(month(dat$LogTime), levels = as.character(1:12))
  dat$WeekDay <- as.factor(weekdays(dat$LogTime, abbreviate = T))
  dat$Day <- as.factor(day(dat$LogTime))
  dat$Hour <- as.factor(hour(dat$LogTime))
  return(dat)
}

rxDataStep(Pi2, outFile = "Line2_Validate1.xdf", overwrite = T, transformFunc = create_DatetimeVars,
           transformPackages = "lubridate", varsToDrop = c("LogMonth", "LogYear"))
Pi2 <- change_Xdf("Line2_Validate.xdf", n = 1)

Pi2 <- Pi2 %>% arrange(Seg, Site, LogTime)
persist(Pi2, "Line2_Validate.xdf")
Pi2 <- RxXdfData("Line2_Validate.xdf")

# Create Differential Column
rxDataStep(Pi2, outFile = "Line2_Validate1.xdf", overwrite = T,
           transforms = list(differential = Discharge_Pressure - Suction_Pressure))
rxDataStep("Line2_Validate1.xdf", outFile = "Line2_Validate2.xdf",
           varsToDrop = c("Discharge_Pressure", "Suction_Pressure"))
Line2 <- change_Xdf("Line2_Validate.xdf", n = 2)

# Rename Columns
plyr::rename(Pi2, c("LogTime" = "ts_pi", "Seg" = "line_segment", "Month" = "ts_month", "Flow_Rate" = "flow_rate",
                    "DRA" = "dra", "HHPU" = "hhpu", "KW" = "kw", "NumPumpsOn" = "pump_status_on", "Hour" = "ts_hour",
                    "TotalPumps" = "pump_total", "WeekDay" = "ts_dow", "Day" = "ts_day", "Site" = "pump_station",
                    "differential" = "differential"))

```

Time to make predictions. See part 5 for more details on everything.

Load Required user defined functions for making and evaluating predictions

```{r Load UDFs}

# Function used to create evaluation statistics for hhpu models
Stats_Calculator <- function(xdf){
  
  xdf <- mutate(xdf, residuals = hhpu - hhpu_Pred)
  hhpu_mean <- mean(xdf$hhpu)
  SSres <- sum(xdf$residuals^2)
  SStot <- sum((xdf$hhpu - hhpu_mean)^2)
  p <- length(names(xdf)) - 2
  n <- nrow(xdf)
  
  MAE <- sum(abs(xdf$residuals))/n
  RMSE <- sqrt(sum(xdf$residuals^2)/n)
  Rsquared <- 1 - SSres/SStot
  Rsquared_adj <- Rsquared - (1 - Rsquared)*(p/(n-p-1))
  
  return(c(MAE,RMSE,Rsquared,Rsquared_adj))
  
}

# Function used calculates the expected number of pumps on for each row of data.
predict_PumpsOn <- function(dat){
  dat$pump_status_on_Pred <- round(0*dat$prob_0 + 1*dat$prob_1 + 2*dat$prob_2 + 3*dat$prob_3 + 
                               4*dat$prob_4 + 5*dat$prob_5 + 6*dat$prob_6)
  dat$pump_status_on_Pred <- factor(dat$pump_status_on_Pred, levels = c(as.character(0:6)))
  return(dat)
}

# Function used to return evaluation statistics for pump_status_on models
Evaluation_Classification_Stats_Calc <- function(xdf){
  
  xdf <- rxDataStep(xdf, outFile = "df.xdf", transformVars = c("pump_status_on", "pump_status_on_Pred"),
                    transforms = list(correct = ifelse(pump_status_on == pump_status_on_Pred, 1, 0)), overwrite = T)
  acc <- sum(xdf$correct)/nrow(xdf)*100
  
  conf_mat <- rxCrossTabs(~pump_status_on_Pred:pump_status_on, xdf)
  conf_mat <- conf_mat$counts$`pump_status_on_Pred:pump_status_on`
  Precision <- diag(conf_mat)/rowSums(conf_mat)
  Recall <- diag(conf_mat)/colSums(conf_mat)
  F1 <- 2*Precision*Recall/(Precision+Recall)
  Precision_Recall <- data.frame("Pumps_On" = c(0:6), Precision, Recall, F1)
  
  file.remove("df.xdf")
  return(list("Accuracy" = acc, "Confusion_Matrix" = conf_mat,"Precision_Recall" = Precision_Recall))

}

```

Let's first predict hhpu using actual pump status on values.

```{r Predict hhpu with pump status actuals}

# Load hhpu models
load("trained_models_NoSch.Rdata")
lm_model <- trained_models$LinMod
glm_model <- trained_models$Glm
dt_model <- trained_models$DTree
df_model <- trained_models$DForest
bdt_model <- trained_models$BTree

# Save the feature we want to keep for matching later on
features <- c("ts_pi", "line_segment", "pump_station", "hhpu")

# Linear Regression
lm_scored <- rxPredict(lm_model, Line2, extraVarsToWrite = features, outData = "Line2_Val_hhpu_lm_scored.xdf",
                       overwrite = T)

lm_Stats <- Stats_Calculator(lm_scored)
MAE_lm <- lm_Stats[1]
RMSE_lm <- lm_Stats[2]
Rsquared_lm <- lm_Stats[3]
Rsquared_adj_lm <- lm_Stats[4]

# General Linear Models
glm_scored <- rxPredict(glm_model, Line2, outData = "Line2_Val_hhpu_glm_scored.xdf", extraVarsToWrite = features,
                        overwrite = T)

glm_Stats <- Stats_Calculator(glm_scored)
MAE_glm <- glm_Stats[1]
RMSE_glm <- glm_Stats[2]
Rsquared_glm <- glm_Stats[3]
Rsquared_adj_glm <- glm_Stats[4]

# Decision Tree
dt_scored <- rxPredict(dt_model, Line2, extraVarsToWrite = features, outData = "Line2_Val_hhpu_dt_scored.xdf",
                       overwrite = T)

dt_Stats <- Stats_Calculator(dt_scored)
MAE_dt <- dt_Stats[1]
RMSE_dt <- dt_Stats[2]
Rsquared_dt <- dt_Stats[3]
Rsquared_adj_dt <- dt_Stats[4]

# Decision Forest
df_scored <- rxPredict(df_model, Line2, extraVarsToWrite = features, outData = "Line2_Val_hhpu_df_scored.xdf",
                       overwrite = T)

df_Stats <- Stats_Calculator(df_scored)
MAE_df <- df_Stats[1]
RMSE_df <- df_Stats[2]
Rsquared_df <- df_Stats[3]
Rsquared_adj_df <- df_Stats[4]

# Boosted Tree
bdt_scored <- rxPredict(bdt_model, Line2, extraVarsToWrite = features, outData = "Line2_Val_hhpu_bdt_scored.xdf",
                        overwrite = T)

bdt_Stats <- Stats_Calculator(bdt_scored)
MAE_bdt <- bdt_Stats[1]
RMSE_bdt <- bdt_Stats[2]
Rsquared_bdt <- bdt_Stats[3]
Rsquared_adj_bdt <- bdt_Stats[4]

# Calculate Evaluation Statistics for hhpu predictions
models <- c("lm", "glm", "dt", "df", "bdt")
MAE <- c(MAE_lm, MAE_glm, MAE_dt, MAE_df, MAE_bdt)
RMSE <- c(RMSE_lm, RMSE_glm, RMSE_dt, RMSE_df, RMSE_bdt)
Rsquared <- c(Rsquared_lm, Rsquared_glm, Rsquared_dt, Rsquared_df, Rsquared_bdt)
Rsquared_adj <- c(Rsquared_adj_lm, Rsquared_adj_glm, Rsquared_adj_dt, Rsquared_adj_df, Rsquared_adj_bdt)

model_stats <- data.frame(models,MAE,RMSE,Rsquared,Rsquared_adj)
write.csv(model_stats, file = "model_stats_Val_hhpu.csv", row.names = F)
print(model_stats)

# Remove models from memory once you're done with them
remove(trained_models)
remove(lm_model)
remove(glm_model)
remove(dt_model)
remove(df_model)
remove(bdt_model)

# Rename prediction columns in our scored xdf data files
plyr::rename(lm_scored, c("hhpu_Pred" = "hhpu_lm_Pred"))
plyr::rename(glm_scored, c("hhpu_Pred" = "hhpu_glm_Pred"))
plyr::rename(dt_scored, c("hhpu_Pred" = "hhpu_dt_Pred"))
plyr::rename(df_scored, c("hhpu_Pred" = "hhpu_df_Pred"))
plyr::rename(bdt_scored, c("hhpu_Pred" = "hhpu_bdt_Pred"))

# Merge the predictions back onto the original data set for easy analysis
rxMerge(inData1 = Line2, inData2 = lm_scored, outFile = "Line2_Validate_scored.xdf",
        type = "left", matchVars = features)
rxMerge(inData1 = "Line2_Validate_scored.xdf", inData2 = glm_scored, outFile = "Line2_Validate_scored1.xdf",
        type = "left", matchVars = features)
rxMerge(inData1 = "Line2_Validate_scored1.xdf", inData2 = dt_scored, outFile = "Line2_Validate_scored2.xdf",
        type = "left", matchVars = features)
rxMerge(inData1 = "Line2_Validate_scored2.xdf", inData2 = df_scored, outFile = "Line2_Validate_scored3.xdf",
        type = "left", matchVars = features)
rxMerge(inData1 = "Line2_Validate_scored3.xdf", inData2 = bdt_scored, outFile = "Line2_Validate_scored4.xdf",
        type = "left", matchVars = features)
Line2 <- change_Xdf("Line2_Validate_scored.xdf", n = 4)

```

Let's graph time series of hhpu vs hhpu_Pred of each model for each station by month. To keep things organized we'll split it by segment.

Each chunk will represent one segment. The the color of a line is:
* black := hhpu actual
* red := hhpu LinMod
* blue := hhpu Glm
* green := hhpu DTree
* yellow := hhpu DForest
* pink := hhpu BTree

2-EP-YP

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-EP-YP"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```

2-YP-KB

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-YP-KB"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```

2-KB-SB

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-KB-SB"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```

2-QU-CM

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-QU-CM"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```

2-CM-GF

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-CM-GF"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```

2-GF-CR

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-GF-CR"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```

2-CR-PR

```{r hhpu vs. hhpu_Pred}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-CR-PR"))

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_lm_Pred), color = "red") + 
    facet_wrap(~pump_station) + ggtitle(paste0("LinMod ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_glm_Pred), color = "blue") + 
    facet_wrap(~pump_station) + ggtitle(paste0("Glm ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_dt_Pred), color = "green") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_df_Pred), color = "yellow") + 
    facet_wrap(~pump_station) + ggtitle(paste0("DForest ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

for(i in 1:2){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) + geom_line(aes(y = hhpu_bdt_Pred), color = "pink") + 
    facet_wrap(~pump_station) + ggtitle(paste0("BTree ",as.character(month(i, label = T, abbr = F))))
  plot(p)
}

```
