# Part 3c - Data Exploration_Sampled

Required Packages

```{r Required Packages}

library(dplyrXdf)
library(ggplot2)
library(lubridate)

```

Data which needs to be load.

```{r Load Data}

Line2 <- RxXdfData("Pi_Schedule.xdf")

```

Since we have so many data points, we need to take a sample of the data if we want to do any scatter plots, etc. For consistency we need to set a seed.

```{r Sample Data}

sample_df <- function(df=Line2_df,st='pump_station',p=0.01){
  
  str <- st
  sdf <- df[0,]
  stu <- unique(df[str])
  len <- nrow(stu)
  for (i in 1:len) {
    
    dfx <- subset(df,df[,str]==stu[i,1])
    
    tr <- floor(nrow(dfx)*p)
    te <- nrow(dfx) - tr
    
    s <- sample(c(rep(1,tr),rep(0,te)))
    sdfx <- subset(dfx,s==1)
    
    sdf <- rbind(sdf,sdfx)
    
  }
  
  return(sdf)
  
}

set.seed(123)
Line2_df <- rxImport(Line2)
Line2_sdf <- sample_df(Line2_df)
remove(Line2_df)

```

Let's look at a scatter plot of HHPU vs. Flow

```{r HHPU vs Flow}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point()

```

We can see a parabolic relationship between flow_rate and hhpu.

Let's filter by segment and station.

```{r HHPU vs. Flow by Segment and Station}

segment <- Line2_sdf %>% select(line_segment) %>% distinct
segment <- as.character(segment[[1]])

for (i in 1:length(segment)){
  p <- ggplot(subset(Line2_sdf, line_segment == segment[i]), aes(x = flow_rate, y = hhpu)) + geom_point() +
    facet_wrap(~pump_station) + ggtitle(segment[i])
  plot(p)
}

```

For some station a linear relationship between flow and hhpu exists, for others a more parabolic relationsip appears to exist.

Let's do the same thing but use color for the stations instead of splitting into different graphs

```{r HHPU vs. Flow by Segment and Station Color}

segment <- Line2_sdf %>% select(line_segment) %>% distinct
segment <- as.character(segment[[1]])

for (i in 1:length(segment)){
  p <- ggplot(subset(Line2_sdf, line_segment == segment[i]), aes(x = flow_rate, y = hhpu, color = pump_station)) + geom_point()
  plot(p)
}

```

The relationship between flow and station seems to change for some of the stations. Some of the stations seem to have a parabolic relationship between flow and hhpu, others linear. There little differences between the stations with a parabolic relationship and between stations with linear.

Let's look at hhpu vs flow for pump_status_on

```{r HHPU vs Flow by pump_status_on}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~pump_status_on)

```

Let's use color this time to look at flow vs hhpu by pump_status_on

```{r Flow vs HHPU by pump_status_on Color}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = pump_status_on)) + geom_point()

```

Clearly the num_pump_status_on matters for the relationship between flow_rate and hhpu

Let's look at flow vs. hhpu by total_pumps

```{r Flow vs. HHPU by total_pumps}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~pump_total)
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = pump_total)) + geom_point()

```

The relationship between flow and hhpu is linear for all values total pumps except for 5. 5 is parabolic. The max flow_rate value for pumps of 2 or less is 2400 which is probably becaues we can't go above that.

Let's look at the relationship by commodity group

```{r Flow vs HHPU by Commodity Group}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~commodity_group)
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = commodity_group)) + geom_point()

```

The relationship between flow and hhpu doesn't seem to change for the commodity groups.

Let's look at how the datetime features affect the relationship between hhpu and flow by using facet

```{r Flow vs HHPU by Datetime Facet}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~ts_day)
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~ts_dow)
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~ts_hour)
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu)) + geom_point() + facet_wrap(~ts_month)

```

and by color for each datetime feature

```{r Flow vs HHPU by Datetime Color}

ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = ts_day)) + geom_point()
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = ts_dow)) + geom_point()
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = ts_hour)) + geom_point()
ggplot(Line2_sdf, aes(x = flow_rate, y = hhpu, color = ts_month)) + geom_point()

```

There doesn't seem to be any changes in the relationship between flow and hhpu for any of the datetime features.

Let's look at some time series now. Will do this for each segment and each month.
Segment "2-CM-CR"

```{r Times Series 2-CM-CR}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-CM-CR"))
sdf$pump_status_on <- (as.numeric(sdf$pump_status_on) - 1)*1000
sdf$ts_year <- year(sdf$ts_pi)
sdf <- subset(sdf,ts_year == 2016)

for (i in 1:12){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf1, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) +
    geom_line(aes(y = pump_status_on), color = "blue") +
    geom_line(aes(y = flow_rate), color = "red") + facet_wrap(~pump_station) +
    ggtitle(as.character(month(i, label = T, abbr = F)))
  plot(p)
}

```

Station CM in August looks to be an issue

Segment "2-CR-PR"

```{r Times Series 2-CR-PR}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-CR-PR"))
sdf$pump_status_on <- (as.numeric(sdf$pump_status_on) - 1)*1000
sdf$ts_year <- year(sdf$ts_pi)
sdf <- subset(sdf,ts_year == 2016)

for (i in 1:12){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf1, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) +
    geom_line(aes(y = pump_status_on), color = "blue") +
    geom_line(aes(y = flow_rate), color = "red") + facet_wrap(~pump_station) +
    ggtitle(as.character(month(i, label = T, abbr = F)))
  plot(p)
}

```

Segment "2-EP-YP"

```{r Times Series 2-KB-SB}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-EP-YP"))
sdf$pump_status_on <- (as.numeric(sdf$pump_status_on) - 1)*1000
sdf$ts_year <- year(sdf$ts_pi)
sdf <- subset(sdf,ts_year == 2016)

for (i in 1:12){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf1, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) +
    geom_line(aes(y = pump_status_on), color = "blue") +
    geom_line(aes(y = flow_rate), color = "red") + facet_wrap(~pump_station) +
    ggtitle(as.character(month(i, label = T, abbr = F)))
  plot(p)
}

```

Segment "2-KB-SB"

```{r Times Series 2-KB-SB}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-KB-SB"))
sdf$pump_status_on <- (as.numeric(sdf$pump_status_on) - 1)*1000
sdf$ts_year <- year(sdf$ts_pi)
sdf <- subset(sdf,ts_year == 2016)

for (i in 1:12){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf1, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) +
    geom_line(aes(y = pump_status_on), color = "blue") +
    geom_line(aes(y = flow_rate), color = "red") + facet_wrap(~pump_station) +
    ggtitle(as.character(month(i, label = T, abbr = F)))
  plot(p)
}

```

Station MI in July is an issue.

Segment "2-QU-CM"

```{r Times Series 2-QU-CM}

sdf <- rxImport(Line2, rowSelection = (line_segment == "2-QU-CM"))
sdf$pump_status_on <- (as.numeric(sdf$pump_status_on) - 1)*1000
sdf$ts_year <- year(sdf$ts_pi)
sdf <- subset(sdf,ts_year == 2016)

for (i in 1:12){
  sdf1 <- subset(sdf, ts_month == i)
  p <- ggplot(sdf1, aes(x = ts_pi)) + geom_line(aes(y = hhpu)) +
    geom_line(aes(y = pump_status_on), color = "blue") +
    geom_line(aes(y = flow_rate), color = "red") + facet_wrap(~pump_station) +
    ggtitle(as.character(month(i, label = T, abbr = F)))
  plot(p)
}

```

Using Information Gain, let's determine which features work are best for predicting HHPU

```{r HHPU Information Gain}

library(FSelector)
Line2_sdf1 <- Line2_sdf[-c(1,8,14,17,18)]
corr <- information.gain(hhpu~.,Line2_sdf1)
col_names <- names(Line2_sdf1)
col_names <- col_names[-c(6)]
corr <- data.frame("Features" = col_names, "Information_Gain" = corr$attr_importance)
corr <- corr[order(corr$Information_Gain, decreasing = T),]
print(corr)

```