# Part 3b - Data Exploration 1: Label Exploration

Required Packages

```{r Required Packages, echo=FALSE, warning=FALSE}

library(dplyrXdf)
library(ggplot2)

```

Data to be loaded

```{r Load Data}

Line2 <- RxXdfData("Pi_Schedule.xdf")

```

Let's turn our attention to our label hhpu

Distribution of HHPU

```{r HHPU Distribution}

rxHistogram(~hhpu,Line2)

```

A lot of hhpu = 0 points. Let's calculate the percentage of data points sucht that hhpu = 0

```{r HHPU = 0 Porportion}

nhhpu0 <- Line2 %>% filter(hhpu == 0) %>% nrow
phhpu0 <- nhhpu0/nrow(Line2)*100
cat("the percentage of points where hhpu = 0 is ", round(phhpu0,2),"%",sep="")

```

Let's look at the dist of hhpu by pump_station

```{r HHPU dist by NumPumpsOn, echo=FALSE}

rxHistogram(~hhpu|pump_station,Line2)

```

Can't see much so let's look at hhpu by segment

```{r HHPU by Segment}

rxHistogram(~hhpu|line_segment,Line2)

```

Let's look at the distribution of hhpu = 0 for each segment

```{r HHPU Distribution by Segment and Station}

rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-CM-CR"), title = "2-CM-CR")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-CR-PR"), title = "2-CR-PR")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-KB-SB"), title = "2-KB-SB")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-QU-CM"), title = "2-QU-CM")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-EP-YP"), title = "2-EP-YP")
rxHistogram(~hhpu|pump_station, Line2, rowSelection = (line_segment == "2-YP-KB"), title = "2-YP-KB")

```

Segments 2-CM-CR and 2-CR-PR had Plenty of points where HHPU didn't equal 0. For the other 4 segments, the majority of points had hhpu = 0.

Let's calculate the percentage of points where hhpu = 0 for each station.

```{r Porportion of Points HHPU = 0 by segment, echo=FALSE, r Porportion of Points HHPU=0 by segment}

hhpu0 <- rxDataStep(Line2, outFile = "Line2_hhpu0.xdf", rowSelection = (hhpu == 0), overwrite = T)

hhpu0_len <- rxCube(~line_segment, hhpu0)
hhpu0_len <- hhpu0_len$Counts
hhpu_len <- rxCube(~line_segment, Line2)
hhpu_len <- hhpu_len$Counts
phhpu0 <- hhpu0_len/hhpu_len*100

segment <- Line2 %>% select(line_segment) %>% distinct %>% head(10)
print(data.frame("Segment" = segment, "Porportion" = phhpu0))

```

For segments 2-EP-YP,, 2-KB-SB, 2-QU-CM, and 2-YP-KB the porportion of points such that hhpu = 0 is approximately half.

Let's look at hhpu by pump_status_on

```{r HHPU by pump_status_on}

rxHistogram(~hhpu|pump_status_on,Line2)

```

Calculate the percentage of data points where NumPumpsOn = 0 given hhpu = 0

```{r pumps_status_on = 0 given hhpu = 0}

hhpu0n0 <- hhpu0 %>% filter(pump_status_on == 0) %>% nrow()
phhpu0n0 <- hhpu0n0/nrow(hhpu0)*100
cat("the percentage of points where pump_status_on = 0 given hhpu = 0 is ",round(phhpu0n0,2),"%",sep="")

```

Given hhpu = 0, every pump is off for about 99.44% of the pumps. Previously we found that 99.3% of points had flow = 0 given all the pumps were off.

Let's look at the percentage of data points where hhpu = 0 given num_pumps_on = 0

```{r Percenteage and Dist hhpu = 0 given pump_status_on = 0}

pump0 <- rxDataStep(Line2, outFile = "pump0.xdf", rowSelection = (pump_status_on == 0), overwrite = T)
nhhpu0 <- pump0 %>% filter(hhpu == 0) %>% nrow()
phhpu0 <- nhhpu0/nrow(pump0)*100

cat("the percentage of points where hhpu = 0 given pump_status_on = 0 is ",round(phhpu0,2),"%",sep="")
rxHistogram(~hhpu,pump0)

```

Let's see how often flow = 0 given hhpu = 0.

```{r Percentage flow = 0 given hhpu = 0}

hhpu0f0 <- hhpu0 %>% filter(flow_rate == 0) %>% nrow()
phhpu0f0 <- hhpu0f0/nrow(hhpu0)*100
cat("the percentage of points where flow = 0 given hhpu = 0 is ",round(phhpu0f0,2),"%",sep="")

```

So roughly 70% of the data points have flow = 0 given hhpu = 0. This is probably because it takes some time for the liquid to stop flowing once the pumps are shut off. Let's determine the distribution of flow_rate given hhpu = 0

```{r Flow Distribution given hhpu = 0}

rxHistogram(~flow_rate,hhpu0)

```

Let's look hhpu = 0 given flow = 0 distribution and porportion.

```{r HHPU = 0 given Flow = 0}

flow0 <- rxDataStep(Line2, outFile = "flow0.xdf", rowSelection = (flow_rate == 0), overwrite = T)
flow0h0 <- flow0 %>% select(hhpu) %>% nrow()
pflow0h0 <- flow0h0/nrow(flow0)*100

cat("the percentage of points where hhpu = 0 given flow = 0 is ", round(pflow0h0,2),"%",sep="")
rxHistogram(~hhpu,flow0)

```

given flow = 0, approximately 100% of the points had hhpu = 0.

So while having no pumps on and hhpu = 0 will not always mean flow = 0 (it takes time to slow down), if flow = 0, hhpu and pump_status_on will almost always be 0.

Let's look at commodity group and hhpu

```{r hhpu by commodity group}

rxHistogram(~hhpu|commodity_group,Line2)

```

# Removing HHPU = 0 from the data

We can see that the vast majority of points with hhpu = 0 also had pump_status_on = 0 but because of this we can't see much else so let's filter out the points = 0.

####NOTE TO SELF: perform data exploration on on a data set with hhpu = 0 filtered out

```{r HHPU != 0 by pump_status}

hhpu_n0 <- rxDataStep(Line2, outFile = "Line2_hhpu_not0.xdf", rowSelection = (hhpu != 0), overwrite = T)

rxHistogram(~hhpu,hhpu_n0)
rxHistogram(~hhpu|pump_status_on,hhpu_n0)

```

Clearly the number of pumps on increases the hhpu value.

Let's look at hhpu and commodity id with hhpu = 0 filtered out.

Let's look at commodity group and hhpu

```{r hhpu by commodity group}

rxHistogram(~hhpu|commodity_group, hhpu_n0)

```

Doesn't seem to be much of a difference in hhpu for each commodity group.


## HHPU vs. Time variables

hhpu by day, day of the week, hour, and month

```{r HHPU by Datetime Distribution}

rxHistogram(~hhpu|ts_day, Line2, title = "Day")
rxHistogram(~hhpu|ts_dow, Line2, title = "Day of Week")
rxHistogram(~hhpu|ts_hour, Line2, title = "Hour")
rxHistogram(~hhpu|ts_month, Line2 , title = "Month")

```

HHPU = 0 is killing this so let's remove the data points where hhpu = 0

```{r HHPU != 0 by Datetime Distribution}

rxHistogram(~hhpu|ts_day, hhpu_n0, title = "Day")
rxHistogram(~hhpu|ts_dow, hhpu_n0, title = "Day of Week")
rxHistogram(~hhpu|ts_hour, hhpu_n0, title = "Hour")
rxHistogram(~hhpu|ts_month, hhpu_n0, title = "Month")

```

HHPU doesn't seem to depend on any of these datetime values.

