# Part 3a - Data Exploration 1: Feature Exploration

Packages required to run this script.

```{r Required Packages, message=FALSE}

library(dplyrXdf)

```

Data needed to run this script.

```{r Data Load, message=FALSE}

Line2 <- RxXdfData("Pi_Schedule.xdf")

```

## Stations on Line 2

The stations found on line 2

```{r Stations, message=FALSE}

station <- Line2 %>% select(pump_station) %>% distinct %>% head(30)
station <- station[[1]]
print(station)

```

28 stations on line 2 (actually 30 but stations SB and IV (?) where lost when we removed Na pi tags).

Number of points from each station

```{r Station Porportions, message=FALSE}

station_len <- rxCrossTabs(~pump_station, Line2)
station_len <- as.vector(station_len$counts$pump_station)
station_prop <- station_len/nrow(Line2)*100

station_size <- data.frame("Station" = station, "Number of Points" = station_len, "Proportion of Points" = station_prop)
print(station_size)

```

Each station takes up roughly 3-5% of the data points except for station CC which only takes up 0.6% of the data.
##### Look into station CC

## Segments on Line 2

The stations found on line 2

```{r Segments, message=FALSE}

segment <- Line2 %>% select(line_segment) %>% distinct %>% head(30)
segment <- segment[[1]]
print(segment)

```

6 segments on line 2.

Number of points from each segment

```{r Segment Porportions, message=FALSE}

segment_len <- rxCrossTabs(~line_segment, Line2)
segment_len <- as.vector(segment_len$counts$line_segment)
segment_prop <- segment_len/nrow(Line2)*100

segment_size <- data.frame("Segment" = segment, "Number of Points" = segment_len, "Proportion of Points" = segment_prop)
print(segment_size)

```

Segments 2-CM-CR take up a little more than half the total amount of data points. Segment 2-YP-KB takes up less than 10% of the points. The other 3 take up around 12-15% of the points each.

What stations are in each segment.

```{r Stations in Segment, message=FALSE}

st_CM_CR <- Line2 %>% filter(line_segment == "2-CM-CR") %>% select(pump_station) %>% distinct()
st_CM_CR <- as.character(st_CM_CR[[1]])
st_CR_PR <- Line2 %>% filter(line_segment == "2-CR-PR") %>% select(pump_station) %>% distinct()
st_CR_PR <- as.character(st_CR_PR[[1]])
st_EP_YP <- Line2 %>% filter(line_segment == "2-EP-YP") %>% select(pump_station) %>% distinct()
st_EP_YP <- as.character(st_EP_YP[[1]])
st_KB_SB <- Line2 %>% filter(line_segment == "2-KB-SB") %>% select(pump_station) %>% distinct()
st_KB_SB <- as.character(st_KB_SB[[1]])
st_QU_CM <- Line2 %>% filter(line_segment == "2-QU-CM") %>% select(pump_station) %>% distinct()
st_QU_CM <- as.character(st_QU_CM[[1]])
st_YP_KB <- Line2 %>% filter(line_segment == "2-YP-KB") %>% select(pump_station) %>% distinct()
st_YP_KB <- as.character(st_YP_KB[[1]])

station_segment <- list("2-CM-CR" = st_CM_CR, "2-CR-PR" = st_CR_PR, "2-EP-YP" = st_EP_YP, "2-KB-SB" = st_KB_SB,
                        "2-QU-CM" = st_QU_CM, "2-YP-KB" = st_YP_KB)
print(station_segment)

```

Looking at the stations in each segment, the reason why 2-YP-KB is less than 10% is that it has station CC which lost a lot of points.

Let's look at the count matrix between segment and station

```{r Count btw Station and Segment, message = FALSE}

rxCrossTabs(~pump_station+line_segment, Line2)

```

The only thing that stands out is station CC once again.

## Activities

The activities found on line 2

```{r Activities, message=FALSE}

activity <- Line2 %>% select(activity_code) %>% distinct %>% head(30)
activity <- activity[[1]]
print(activity)

```

3 acctivities on line 2 as well some NA values.

Number of points from each activity

```{r Activity Porportions, message=FALSE}

activityl <- levels(activity)

activity_len <- rxCrossTabs(~activity_code, Line2)
activity_len <- as.vector(activity_len$counts$activity_code)
activity_prop <- activity_len/nrow(Line2)*100
s <- sum(activity_len)
Na <- nrow(Line2) - s
activityl <- c(activityl, "NA")
activity_len <- c(activity_len, Na)
activity_prop <- c(activity_prop, Na/nrow(Line2)*100)

activity_size <- data.frame("Activity" = activityl, "Number of Points" = activity_len,
                            "Proportion of Points" = activity_prop)
print(activity_size)

```

60% of the data points are evens, 9% are injections, and 1% are receipts. The rest are NA values created because their was no matching schedule data for the pi tags. No landings or deliveries made it into the data.

## Commodity Group

The commodity groups pushed through line 2

```{r Activities, echo=TRUE, message=FALSE, warning=FALSE}

commodity <- Line2 %>% select(commodity_group) %>% distinct %>% head(30)
commodity <- commodity[[1]]
print(commodity)

```

5 commodity groups are pushed on line 2 as well some NA values.

Number of points from each commodity

```{r Commodity Porportions, echo=TRUE, message=FALSE, warning=FALSE}

commodityl <- levels(commodity)

commodity_len <- rxCrossTabs(~commodity_group, Line2)
commodity_len <- as.vector(commodity_len$counts$commodity_group)
commodity_prop <- commodity_len/nrow(Line2)*100
s <- sum(commodity_len)
Na <- nrow(Line2) - s
commodityl <- c(commodityl, "NA")
commodity_len <- c(commodity_len, Na)
commodity_prop <- c(commodity_prop, Na/nrow(Line2)*100)

commodity_size <- data.frame("Commodity Group" = commodityl, "Number of Points" = commodity_len,
                            "Proportion of Points" = commodity_prop)
print(commodity_size)

```

The majority of commodities pushed on Line 2 belong to the groups LightSweet and LightSynthetic. Condensate and LightSour are pushed as well occasionally. Once again 30% of the data points are Na's due to the join between schedule and pi data.

```{r Commodity by Station and Segment Count, echo=FALSE, warning=FALSE}

rxHistogram(~commodity_group|pump_station,Line2)
rxHistogram(~commodity_group|line_segment,Line2)

```

Due to the issues with merging the schedule and pi data, analyzing the commodity group and activity might not be a god idea yet as it probably isn't accurate.

## Flow_Rate

Let's explore the distribution of flow_rate

```{r Flow_Rate Distribution, echo=TRUE, message=FALSE}

rxHistogram(~ flow_rate, Line2)

```

3 Hills stand out. The largest is at 0 which says that there was nothing flowing along line 2 for a significant portion of time. Their are smaller humps around flow rate 1500 and 3000.

Let's filter by station now

```{r Flow by Station, echo=TRUE, message=FALSE}

rxHistogram(~ flow_rate | pump_station, Line2)

```

Of the 28 graphs, there are two patterns. Either the graphs have two humps at 0 and 1500 or 1 hump at 3000.

Let's filter by segment now

```{r Flow by Segment, echo=TRUE, message=FALSE}

rxHistogram(~ flow_rate | line_segment, Line2)

```

Segments 2-CM-CR and 2-CR-PR have humps around 3000 with few points flow_rate = 0. Segment 2-YP-KB have almost all points flow_rate = 0. The others seem to match the other graph type with humps at 0, 1500, and 3000.

Let's compute the percentage of data points for Line 2 where flow_rate = 0

```{r Flow = 0, message = FALSE}

Line2_flow0 <- rxDataStep(Line2, outFile = "Line2_flow0.xdf", rowSelection = (flow_rate == 0) , overwrite = T)
n_flow0 <- nrow(Line2_flow0)
p_flow0 <- n_flow0/nrow(Line2)*100
cat("The percentage of points where flow = 0 is ", round(p_flow0,2), "%", sep = "")

```

Roughly 20% of the data has flow_rate = 0%

Let's compute this for each station

```{r Flow = 0 by Station, echo=TRUE, message=FALSE, warning=FALSE}

station_len0 <- rxCrossTabs(~pump_station, Line2_flow0)
station_len0 <- as.vector(station_len0$counts$pump_station)
station_prop0 <- station_len0/station_len*100

station_size <- data.frame("Station" = station, "Number of Points" = station_len0, "Porportion" = station_prop0)
print(station_size)

```

Some stations the percentage of data points where flow_rate = 0 is between 2-6%. Some stations the percentage of points is between 20-30%. For stations CC, ME, and YP the percentage is between 45-48%.
######NOTE: look into stations CC, ME, YP

##Number of pumps on

Let's look at the distribution of pumps


```{r Pump_status_on Counts, echo=TRUE, message=FALSE, warning=FALSE}

pump <- Line2 %>% select(pump_status_on) %>% distinct %>% head(10)
pump <- as.character(pump[[1]])
pump <- c(pump,6)
pump_len <- rxCrossTabs(~pump_status_on, Line2)
pump_len <- as.vector(pump_len$counts$pump_status_on)
pump_prop <- round(pump_len/nrow(Line2)*100,2)

pump_size <- data.frame("Pump" = pump, "Number of Points" = pump_len, "Proportion of Points" = pump_prop)
print(pump_size)

```

```{r Pumps On Count by Station, echo=TRUE, message=FALSE}

rxCrossTabs(~pump_status_on:pump_station,Line2)

```

```{r Pumps On Distribution by Station, echo=TRUE, message=FALSE}

rxHistogram(~pump_status_on|pump_station,Line2)

```

```{r PumpsOn Counts by Segment, echo=TRUE}

rxCrossTabs(~pump_status_on:line_segment,Line2)

```

```{r PumpsOn by Segment, echo=TRUE, message=FALSE, warning=FALSE}

rxHistogram(~pump_status_on|line_segment,Line2)

```

```{r NumPumpsOn Counts by segment and station, echo=TRUE, message=FALSE, warning=FALSE}

rxCube(~line_segment:pump_station:pump_status_on, Line2, removeZeroCounts = T )

```


```{r NumPumpsOn Distribution by Segment and Station, echo=TRUE, message=FALSE, warning=FALSE}

rxHistogram(~pump_status_on|pump_station, Line2, rowSelection = (line_segment==seg[1]),
            transformObjects = list(seg = segment), title = "2-CM-CR")
rxHistogram(~pump_status_on|pump_station, Line2, rowSelection = (line_segment==seg[2]),
            transformObjects = list(seg = segment), title = "2-CR-PR")
rxHistogram(~pump_status_on|pump_station, Line2, rowSelection = (line_segment==seg[3]),
            transformObjects = list(seg = segment), title = "2-EP-YP")
rxHistogram(~pump_status_on|pump_station, Line2, rowSelection = (line_segment==seg[4]),
            transformObjects = list(seg = segment), title = "2-KB-SB")
rxHistogram(~pump_status_on|pump_station, Line2, rowSelection = (line_segment==seg[5]),
            transformObjects = list(seg = segment), title = "2-QU-CM")
rxHistogram(~pump_status_on|pump_station, Line2, rowSelection = (line_segment==seg[6]),
            transformObjects = list(seg = segment), title = "2-YP-KB")

```

Let's look at how flow rate is affected by num of pumps on

```{r Flow Distribution by Num Pumps On, message=FALSE}

rxHistogram(~flow_rate|pump_status_on,Line2)

```

It looks like the vast majority of data points where flow = 0 correspond to having pump_status_on = 0 as well. Let's investigate this further.

```{r Flow = 0 by pump_status_on, echo=TRUE, message=FALSE, warning=FALSE}

rxHistogram(~flow_rate|pump_status_on,Line2_flow0)

```

```{r Flow = 0 by pump_status_on Counts, echo=TRUE, message=FALSE, warning=FALSE}

pump <- Line2_flow0 %>% select(pump_status_on) %>% distinct %>% head(10)
pump <- as.character(pump[[1]])
pump <- c(pump,5,6)
pump_len_flow0 <- rxCrossTabs(~pump_status_on, Line2_flow0)
pump_len_flow0 <- as.vector(pump_len_flow0$counts$pump_status_on)
pump_prop_flow0 <- round(pump_len_flow0/nrow(Line2_flow0)*100,2)

pump_size_flow0 <- data.frame("Pump" = pump, "Number of Points" = pump_len_flow0, "Proportion of Points" = pump_prop_flow0)
print(pump_size_flow0)

```

99% of the points where flow = 0, every pump was off. This makes sense.

## Total Pumps

Let's look at the count and distribution of total pumps

```{r Total Pumps Count, echo=TRUE, warning=FALSE}

tpump <- Line2 %>% select(pump_total) %>% distinct %>% head(10)
tpump <- as.character(tpump[[1]])
tpump <- c(tpump,6)
tpump_len <- rxCrossTabs(~pump_total, Line2)
tpump_len <- as.vector(tpump_len$counts$pump_total)
tpump_prop <- round(tpump_len/nrow(Line2)*100,2)

tpump_size <- data.frame("Pump" = tpump, "Number of Points" = tpump_len, "Proportion of Points" = tpump_prop)
print(tpump_size)


```

Let's look at this count by station

```{r Total Pumps by Station and Segment, echo=FALSE, warning=FALSE}

rxCrossTabs(~pump_station:pump_total:line_segment,Line2,removeZeroCounts = T, marginals = T)

```

For every station, approximately 100% of the points all have the same total pump number which makes sense. However for a couple of the stations, a handful (5 or 6) of points state the station has zero points. Let's look further into these rows.

```{r Stations where Total_Pumps = 0, echo=FALSE, warning=FALSE}

tpumps0 <- rxDataStep(Line2, outFile = "Line2_tpumps0.xdf", rowSelection = (pump_total == 0), overwrite = T)
nrow(tpumps0)
tpumps0 %>% select(pump_station) %>% distinct %>% head(30)
tpumps0 %>% filter(pump_station == "CM") %>% head(10)
tpumps0 %>% filter(pump_station == "DN") %>% head(10)
tpumps0 %>% filter(pump_station == "LP") %>% head(10)
tpumps0 %>% filter(pump_station == "KB") %>% head(10)
tpumps0 %>% filter(pump_station == "AP") %>% head(10)

```

######We can either just remove these rows or try to fix them logically by replacing pump_total with the actual pump total for that station and pump_status_on with the previous one.

Let's look at flow by total_pumps

```{r Flow Distribution by Total_Pumps, echo=FALSE, warning=FALSE}

rxHistogram(~flow_rate|pump_total, Line2)

```

Flow and pump_total are most likely not related so let's move on.

Let's look at flow vs. the datetime features

```{r Flow by Datetime Distribution}

rxHistogram(~flow_rate|ts_day, Line2, title = "Day")
rxHistogram(~flow_rate|ts_dow, Line2, title = "Day of Week")
rxHistogram(~flow_rate|ts_hour, Line2, title = "Hour")
rxHistogram(~flow_rate|ts_month, Line2 , title = "Month")

```

There doesn't seem to be any interaction between flow and the datetime values.

##DRA

Distribution of DRA

```{r DRA Distribution, echo=FALSE}

rxHistogram(~dra,Line2)

```

Majority of the data points seem to have dra = 0 but let's confirm

```{r dra = 0 porportion, echo=FALSE, r dra=0 porportion}

dra0 <- Line2 %>% filter(dra == 0)
cat("the percentage of points where dra was 0 is ", round(nrow(dra0)/nrow(Line2)*100,2), "%", sep = "")

```

So for a little more then half the data points, no dra was used. Let's investigate dra further by segment and station.

```{r Dra Distribution by Segment and Station, echo=FALSE, warning=FALSE}

rxHistogram(~dra|pump_station, Line2, rowSelection = (line_segment=="2-CM-CR"), title = "2-CM-CR")
rxHistogram(~dra|pump_station, Line2, rowSelection = (line_segment=="2-CR-PR"), title = "2-CR-PR")
rxHistogram(~dra|pump_station, Line2, rowSelection = (line_segment=="2-EP-YP"), title = "2-EP-YP")
rxHistogram(~dra|pump_station, Line2, rowSelection = (line_segment=="2-KB-SB"), title = "2-KB-SB")
rxHistogram(~dra|pump_station, Line2, rowSelection = (line_segment=="2-QU-CM"), title = "2-QU-CM")
rxHistogram(~dra|pump_station, Line2, rowSelection = (line_segment=="2-YP-KB"), title = "2-YP-KB")


```

It appears that segments 2-CM-CR and 2-CR-PR are the only two segments which use DRA more often than not. The other segments seem to only use it sporatically.

Let's look at dra by commodiy group. DRA might be used for specific commodities.

```{r dra by commodity group, echo=FALSE}

rxHistogram(~dra|commodity_group, Line2)

```

Seems like all the data points with dra were the commodity groups with NA values.

## Datetime features

Let's look at counts of the datetime features just to make sure there aren't any issues.

```{r Datetime}

rxCrossTabs(~ts_day,Line2)
rxCrossTabs(~ts_dow,Line2)
rxCrossTabs(~ts_hour,Line2)
rxCrossTabs(~ts_month,Line2)

```

The counts are all approximately the same so it seems to be okay.

